---
title: "Script Digital Marketing"
author: "Alberto Filosa"
date: "17/8/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r Setup e Librerie, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center",
                      out.width = "60%")

#-- Libraries
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
library(forcats)
library(lubridate)
library(RQuantLib)

set.seed(123456) #-- Set seed for same results

data_dir <- "C:\\Users\\cafil\\Desktop\\Digital-Marketing\\Dataset\\"
```

# Reading Files

## Dataset 1
`raw_1_cli_fid.csv` contains info on the fidelty subscriptions of each costumer account:

* `ID_CLI`: identify client (*Foreign Key*);
* `ID_FID`: identify fidelty program (**Key**);
* `ID_NEG`: identify reference stoe;
* `TYP_CLI_FID`: identify the main account (Binomyal);
* `COD_FID`: identify the fidelty program;
* `STATUS_FID`: identify if an account is active (Binomyal);
* `DT_ACTIVE`: identify the date of activation.

```{r Reading File 1, results='asis'}
#-- Changed name `df_1_cli_fid` in fidelity
fidelity <- read_csv2(file.path(data_dir, "raw_1_cli_fid.csv"), #-- Directory
                          na = c("NA", ""))                     #-- Identify Missing Values

knitr::kable(head(fidelity))    #-- First 5 rows
summary(fidelity)               #-- Descriptive Analysis
```

## Dataset 2
`raw_2_cli_account.csv` contains info on each customer account:

* `ID_CLI`: identify the client (**Key**);
* `EMAIL_PROVIDER`: identify the email account provider;
* `W_PHONE`: identify if a phone number is added (Binomyal);
* `ID_ADDRESS`: identify the address (*Foreign Key*);
* `TYP_CLI_ACCOUNT`: identify the account type of the client;
* `TYP_JOB`: identify the client job.

```{r Reading File 2}
#-- Changed name `df_2_cli_account` in account
account <- read_csv2(file.path(data_dir, "raw_2_cli_account.csv"), #-- Directory
                              na = c("NA", ""))                    #-- Identify Missing Values

knitr::kable(head(account))    #-- First 5 rows
summary(account)               #-- Descriptive Analysis
```

## Dataset 3
`raw_3_cli_address.csv` contains information on the address corresponding to a customer account:

* `ID_ADDRESS`: identify the address (**Key**);
* `CAP`: identify the postal code;
* `PRV`: identify the province;
* `REGION`: identify the region.

```{r Reading File 3}
#-- Changed name `df_3_cli_address` in address
address <- read_csv2(file.path(data_dir, "raw_3_cli_address.csv"), #-- Directory
                              na = c(""))                          #-- Identify Missing Values

knitr::kable(head(address))    #-- First 5 rows
summary(address)               #-- Descriptive Analysis
```

## Dataset 4
`raw_4_cli_privacy.csv` contains information on the privacy policies accepted by each customer:

* `ID_CLI`: identify the client (*Foreign Key*);
* `FLAG_PRIVACY_1`: identify the flag privacy (binomyal);
* `FLAG_PRIVACY_2`: identify the flag profiling (*Foreign Key*);
* `FLAG_DIRECT_MKT`: identify the flag direct marketing (binomyal).

```{r Reading File 4}
#-- Changed name `df_4_cli_privacy` in privacy
privacy <- read_csv2(file.path(data_dir, "raw_4_cli_privacy.csv"), #-- Directory
                              na = c("NA", ""))                    #-- Identify Missing Values

knitr::kable(head(privacy))    #-- First 5 rows
summary(privacy)               #-- Descriptive Analysis
```

## Dataset 5
`raw_5_camp_cat.csv` contains the categorization of the marketing email communications:

* `ID_CAMP`: identify the email campaign (**Key**);
* `TYP_CAMP`: identify the type email campaign.

```{r Reading File 5}
#-- Changed name `df_5_camp_cat` in campaign_category
campaign_category <- read_csv2(file.path(data_dir, "raw_5_camp_cat.csv"), #-- Directory
                           na = c("NA", ""))                              #-- Identify Missing Values

knitr::kable(head(campaign_category))    #-- First 5 rows
summary(campaign_category)               #-- Descriptive Analysis
```

## Dataset 6
`raw_6_camp_event.csv` contains the events (sents, opens and clicks) related to the marketing email communications:

* `ID_EVENT`: identify the feedback event (**Key**);
* `ID_CLI`: identify the client (*Foreign Key*);
* `ID_CAMP`: identify the email campaign (*Foreign Key*);
* `ID_DELIVERY`: identify the delivery;
* `TYP_EVENT`: identify the feedback event:
    + S = Send;
    + V = Open;
    + C = Click;
    + B = Bounce;
    + E = Error;
* `EVENT_DATE`: identify the datetime event.

```{r Reading File 6}
#-- Changed name `df_6_camp_event` in campaign_event
campaign_event <- read_csv2(file.path(data_dir, "raw_6_camp_event.csv"), #-- Directory
                             na = c("NA", ""))                           #-- Identify Missing Values

knitr::kable(head(campaign_event))    #-- First 5 rows
summary(campaign_event)               #-- Descriptive Analysis
```

## Dataset 7
`raw_7_tic.csv` contains the purchase and refund transaction of each customer:

* `ID_SCONTRINO`: identify the transaction (all products have same ID);
* `ID_CLI`: identify the client (*Foreign Key*);
* `ID_NEG`: identify the reference store (*Foreign Key*);
* `ID_ARTICOLO`: identify the purchased or refund item;
* `COD_REPARTO`: identify the business unit corresponding to the item;
* `DIREZIONE`: identify the purchase (1) or refund (-1);
* `IMPORTO_LORDO`: identify the gross amount as the sum of net amount and the discount applied;
* `SCONTO`: identify the discount applied (negative if refund);
* `DATETIME`: datetime of the transaction.

```{r Reading File 7}
#-- Changed name `df_7_tic` in tickets
tickets <- read_csv2(file.path(data_dir, "raw_7_tic.csv"), #-- Directory
                      na = c("NA", ""))                    #-- Identify Missing Values
                      
knitr::kable(head(tickets))    #-- First 5 rows
summary(tickets)               #-- Descriptive Analysis
```

# Cleaning Files

First of all, it is reccomended to create a copy for all datasets before data manipulation: 

```{r Copy Files}
fidelity_clean <- fidelity          #-- Dataset 1
account_clean  <- account           #-- Dataset 2
address_clean  <- address           #-- Dataset 3
privacy_clean  <- privacy           #-- Dataset 4
campaign_category_clean <- campaign_category #-- Dataset 5
campaign_event_clean    <- campaign_event    #-- Dataset 6
tickets_clean  <- tickets           #-- Dataset 7
```

## 1. Fidelity

First, we check if there are some duplicate rows:

```{r Duplicates Fidelity}
fidelity_clean_duplicate <- fidelity_clean %>%
                              summarize(TOT_ID_CLIs = n_distinct(ID_CLI), #-- Number of Total Clients
                              TOT_ID_FIDs = n_distinct(ID_FID),           #-- Number of Total Fidelity Program
                              TOT_ID_CLIFIDs = n_distinct(paste0(as.character(ID_CLI), "-",
                                                                 as.character(ID_FID))),
                              TOT_ROWs = n())                             #-- Number of Total Rows

knitr::kable(fidelity_clean_duplicate)
```

```{r Manipulation Fidelity}
fidelity_clean <- fidelity_clean                                 %>% 
                    mutate(DT_ACTIVE = as.Date(DT_ACTIVE))       %>% #-- Formatting dates 
                    mutate(TYP_CLI_FID = as.factor(TYP_CLI_FID)) %>% #-- Factorize TYP_CLI_FID
                    mutate(STATUS_FID = as.factor(STATUS_FID))       #-- Factorize STATUS_FID
```

Check the number of fidelity subscriptions per client:

```{r Fidelity Subscriptions 1, eval = FALSE}
num_fid_x_cli <- fidelity_clean     %>%
                   group_by(ID_CLI) %>%                         #-- Group by Client
                   summarize(NUM_FIDs  = n_distinct(ID_FID),    #-- Distinct Fidelity
                             NUM_DATEs = n_distinct(DT_ACTIVE)) #-- Distinct Date
```


```{r Fidelity Subscriptions 2}
load("fid_cli.RData")

tot_id_cli <- n_distinct(num_fid_x_cli$ID_CLI)

dist_num_fid_x_cli <- num_fid_x_cli %>%
                        group_by(NUM_FIDs, NUM_DATEs)            %>% #-- Group by Client & Date of Activation
                        summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Total Clients
                        mutate(PERCENT_CLIs = TOT_CLIs/tot_id_cli)   #-- Percentage

knitr::kable(dist_num_fid_x_cli)
```

Note that there are clients with multiple fidelity subscriptions. In details, let examine clients with multiple subscriptions:

```{r Multiple Fidelity}
num_fid_x_cli %>%
  filter(NUM_FIDs == 3)

fidelity_clean %>%
  filter(ID_CLI == 621814)

fidelity_clean %>%
  filter(ID_CLI == 320880)
```

After, we combine the information for each client from first subscription (registration date and store for registration) to the last (type of fidelity and status):

```{r Import Info Combined Fidelty, include = FALSE}
load("cli_fid_first_last.RData")
```

```{r Info Combined Fidelity, eval = FALSE}
client_fidelity_first <- fidelity_clean                        %>%
                           group_by(ID_CLI)                    %>% #-- Grouped by Client ID
                           filter(DT_ACTIVE == min(DT_ACTIVE)) %>% #-- Filtered by min DT_ACTIVE
                           arrange(ID_FID)                     %>% #-- Sorted by client ID
                           filter(row_number() == 1)           %>% #-- Filter Row #1
                           ungroup()                           %>% 
                           as.data.frame()

client_fidelity_last <- fidelity_clean                        %>%
                          group_by(ID_CLI)                    %>% #-- Grouped by Client ID
                          filter(DT_ACTIVE == max(DT_ACTIVE)) %>% #-- Filtered by max DT_ACTIVE
                          arrange(desc(ID_FID))               %>% #-- Sorted by client ID
                          filter(row_number() == 1)           %>% #-- Filter Row #1
                          ungroup()                           %>%
                          as.data.frame()
```

We left join `client_fidelity_first` and `num_fid_x_cli` in `df_1_cli_fid_last`: 

```{r Left Join Fidelity}
fidelity_clean <- client_fidelity_last                      %>%
                      select(ID_CLI,
                             ID_FID,
                             LAST_COD_FID = COD_FID,
                             LAST_TYP_CLI_FID = TYP_CLI_FID,
                             LAST_STATUS_FID = STATUS_FID,
                             LAST_DT_ACTIVE = DT_ACTIVE)    %>% #-- Selection of Variables
                      left_join(client_fidelity_first       %>% #-- Left Join with client_fidelity_first
                                  select(ID_CLI,
                                         FIRST_ID_NEG = ID_NEG,
                                         FIRST_DT_ACTIVE = DT_ACTIVE),
                                by = "ID_CLI")              %>%
                      left_join(num_fid_x_cli         %>%       #-- Left Join with num_fid_x_cli
                                  select(ID_CLI,
                                         NUM_FIDs)          %>%
                                  mutate(NUM_FIDs = as.factor(NUM_FIDs)),
                                by = "ID_CLI")
```

### Exploration Analysis
First, it's necessary to compute the distribution of the fidelity program:

```{r Distribution LAST_COD_FID Fidelity}
df1_dist_codfid <- fidelity_clean          %>%
  group_by(LAST_COD_FID)                   %>% #-- Grouped by Fidelty Program
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                       #-- Sort by Percent decreasing

knitr::kable(df1_dist_codfid)

ggplot(data = df1_dist_codfid,
       aes(x = LAST_COD_FID, y = TOT_CLIs)) +  #-- Dataset to Plot
  geom_bar(stat = "identity",                  
           fill = "steelblue") +               #-- Bar Plot
  theme_minimal()                              #-- ggplot Theme
```

Let's explore the status variable (`LAST_STATUS_FID`):

```{r Distribution LAST_STATUS_FID Fidelity}
df1_dist_status <- fidelity_clean          %>% 
  group_by(LAST_STATUS_FID)                %>%   #-- Grouped by Active Account
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>%   #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>%   #-- Percent
  arrange(desc(PERCENT))                         #-- Sort by Percent decreasing


ggplot(data = df1_dist_status, 
       aes(x = LAST_STATUS_FID, y = TOT_CLIs)) + #-- Dataset to Plot
  geom_bar(stat = "identity",                 
           fill = "steelblue") +                 #-- Bar Plot
  theme_minimal()                                #-- ggplot Theme
```

Let's explore the fidelity card variable (`NUM_FIDs`):

```{r Distribution NUM_FIDs Fidelity}
df1_dist_num <- fidelity_clean             %>%
  group_by(NUM_FIDs)                       %>% #-- Grouped by Fidelity Program
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                       #-- Sort by Percent decreasing

ggplot(data = df1_dist_num,               
       aes(x = NUM_FIDs, y = TOT_CLIs)) +      #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +               #-- Bar Plot
  theme_minimal()                              #-- ggplot Theme
```

Let's explore the last fidelity card variable (`LAST_TYP_CLI_FID`):

```{r Distribution LAST_TYP_CLI_FID Fidelity}
df1_dist_type <- fidelity_clean            %>%
  group_by(LAST_TYP_CLI_FID)               %>%    #-- Grouped by Main Account
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>%    #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>%    #-- Percent
  arrange(desc(PERCENT))                          #-- Sort by Percent decreasing

ggplot(data = df1_dist_type,
       aes(x = LAST_TYP_CLI_FID, y = TOT_CLIs)) + #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +                  #-- Bar Plot
  theme_minimal()                                 #-- ggplot Theme
```

Let's explore the date of activation card variable (`FIRST_DT_ACTIVE`):

```{r Distribution FIRST_DT_ACTIVE Fidelity}
df1_dist_date <- fidelity_clean            %>%
  group_by(FIRST_DT_ACTIVE)                %>%   #-- Grouped by Date of Activation
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>%   #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>%   #-- Percent
  arrange(desc(PERCENT))                         #-- Sort by Percent decreasing

ggplot(data = df1_dist_date,
       aes(x = FIRST_DT_ACTIVE, y = TOT_CLIs)) + #-- Dataset to Plot
  geom_line() +                                  #-- Bar Plot
  theme_minimal()                                #-- ggplot Theme
```

## 2. Account

We check for duplicate rows:

```{r Duplicated Account}
account_clean_duplicate <- account_clean %>%
                            summarize(TOT_ID_CLIs = n_distinct(ID_CLI),
                                      TOT_ROWs = n())
                          
knitr::kable(account_clean_duplicate) #-- No Duplicated rows
```

```{r Manipulation Account}
account_clean <- account_clean %>%
                  mutate(W_PHONE = as.factor(W_PHONE))                  %>% #-- Factorized
                  mutate(TYP_CLI_ACCOUNT = as.factor(TYP_CLI_ACCOUNT))  %>% #-- Factorized 
                  mutate(W_PHONE = fct_explicit_na(W_PHONE, "0"))       %>% #-- Missing Values
                  mutate(EMAIL_PROVIDER = fct_explicit_na(EMAIL_PROVIDER,
                                                          "(missing)")) %>% #-- NA with Categorial
                  mutate(TYP_JOB = fct_explicit_na(TYP_JOB, "(missing)"))   #-- NA with Categorial
```

### Reshaping

```{r Email Provider Account}
df_2_dist_emailprovider <- account_clean                             %>%
                            group_by(EMAIL_PROVIDER)                 %>% #-- Group by EMAIL_PROVIDER
                            summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
                            mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
                            arrange(desc(PERCENT))                   %>% #-- Sorted by PERCENt
                            as.data.frame()

knitr::kable(head(df_2_dist_emailprovider))
```

There are `r n_distinct(df_2_dist_emailprovider$EMAIL_PROVIDER)`: too many different values for `EMAIL_PROVIDER` to be an useful category. We keep the most frequent email provider values and add a common factor level `other` for the remaining:

```{r Reshape Email Provider Account}
knitr::kable(df_2_dist_emailprovider                       %>%
  arrange(desc(PERCENT))                                   %>% #-- Sorted by PERCENt
  mutate(PERCENT_COVERED = cumsum(TOT_CLIs)/sum(TOT_CLIs)) %>% #-- Percent
  as.data.frame()                                          %>%
  head(10))

clean_email_providers <- df_2_dist_emailprovider %>%
                          arrange(desc(PERCENT)) %>%                                        #-- Sorted by PERCENt
                          mutate(PERCENT_COVERED = cumsum(TOT_CLIs)/sum(TOT_CLIs)) %>%      #-- Percent
                          mutate(EMAIL_PROVIDER = as.character(EMAIL_PROVIDER))    %>%
                          mutate(AUX = if_else(PERCENT_COVERED < 0.85 |
                                                 (PERCENT_COVERED > 0.85 &
                                                    lag(PERCENT_COVERED) < 0.85), 1,0)) %>% #-- Percent Covered
                          mutate(EMAIL_PROVIDER_CLEAN = if_else(AUX | EMAIL_PROVIDER == "(missing)",
                                                                EMAIL_PROVIDER,
                                                                "others"))                  #-- Others if Missing

knitr::kable(head(clean_email_providers, 7))

account_clean <- account_clean                          %>%
  mutate(EMAIL_PROVIDER = as.character(EMAIL_PROVIDER)) %>% #-- EMAIL_PROVIDER as Character
  left_join(clean_email_providers                       %>%
              select(EMAIL_PROVIDER, EMAIL_PROVIDER_CLEAN),
            by = "EMAIL_PROVIDER")                      %>% #-- Left Join with clean_email_providers
  select(-EMAIL_PROVIDER)                               %>%
  mutate(EMAIL_PROVIDER_CLEAN = as.factor(EMAIL_PROVIDER_CLEAN)) #-- Factorized
```

```{r Phone Distribution Account}
df_2_dist_phone <- account_clean %>%
  group_by(W_PHONE)                        %>% #-- Grouped By W_PHONE
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                   %>% #-- Sorted by PERCENT
  as.data.frame()

knitr::kable(df_2_dist_phone)
```

```{r Type Account Distribution Account}
df_2_dist_account <- account_clean         %>%
  group_by(TYP_CLI_ACCOUNT)                %>% #-- Grouped By TYP_CLI_ACCOUNT
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                   %>% #-- Sorted by PERCENT
  as.data.frame()

knitr::kable(df_2_dist_account)
```

### Exploration Analysis

Let's explore the email provider for each client (`EMAIL_PROVIDER_CLEAN`):

```{r Distribution EMAIL_PROVIDER_CLEAN Account}
df2_dist_emailproviderclean <- account_clean %>%
  group_by(EMAIL_PROVIDER_CLEAN)             %>% #-- Grouped By EMAIL_PROVIDER_CLEAN
  summarize(TOT_CLIs = n_distinct(ID_CLI))   %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs))   %>% #-- Percent
  arrange(desc(PERCENT))                         #-- Sorted by PERCENT 

ggplot(data = df2_dist_emailproviderclean,
       aes(x = EMAIL_PROVIDER_CLEAN,
           y = TOT_CLIs)) +                      #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +                 #-- Bar Plot
  theme_minimal()                                #-- ggplot Theme
```

Let's explore the type job for each client (`TYP_JOB`):

```{r Distribution TYP_JOB Account}
df_2_dist_job <- account_clean             %>%
  group_by(TYP_JOB)                        %>% #-- Grouped By TYP_JOB
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                   %>% #-- Sorted by PERCENT 
  as.data.frame()

knitr::kable(df_2_dist_job)

ggplot(data = df_2_dist_job,
       aes(x = TYP_JOB,
           y = TOT_CLIs)) +                    #-- Dataset to Plot
  geom_bar(stat = "identity", 
           fill = "steelblue") +               #-- Bar Plot
  theme_minimal()                              #-- ggplot Theme
```

Let's explore the type account for each client (`TYP_CLI_ACCOUNT`):

```{r Distribution TYP_CLI_ACCOUNT Account}
df_2_dist_account <- account_clean         %>%
  group_by(TYP_CLI_ACCOUNT)                %>% #-- Grouped By TYP_CLI_ACCOUNT
  summarize(TOT_CLIs = n_distinct(ID_CLI)) %>% #-- Number of Total Client
  mutate(PERCENT = TOT_CLIs/sum(TOT_CLIs)) %>% #-- Percent
  arrange(desc(PERCENT))                   %>% #-- Sorted by PERCENT 
  as.data.frame()

knitr::kable(df_2_dist_account)

ggplot(data = df_2_dist_phone,
       aes(x = W_PHONE,
           y = TOT_CLIs)) +                    #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +               #-- Bar Plot
  theme_minimal()                              #-- ggplot Theme
```

## 3. Address

We check for duplicate rows:

```{r Duplicated Address}
knitr::kable(address_clean %>%
               summarize(TOT_ID_ADDRESSes = n_distinct(ID_ADDRESS), #-- Total of Addresses
               TOT_ROWs = n()))                                     #-- Total Rows

#-- Remove Duplicates
address_clean <- address_clean %>%
                  distinct()
```

Let's manipulate missing values:

```{r Manipulation Address}
address_clean <- address_clean %>%
                    mutate(CAP = as.character(CAP))                   %>% #-- Factorized
                    filter(!is.na(CAP) & !is.na(PRV) & !is.na(REGION))
```

### Explanatory Analysis

Let's explore the CAP address (`CAP`):

```{r Distribution CAP Address}
nrow(address_clean %>%
  distinct(CAP))       #-- Number of CAP
```

There are `r nrow(address_clean %>% distinct(CAP))`: too many to plot.

Let's explore the province address (`PRV`):

```{r Distribution Province Address}
nrow(address_clean %>%
  distinct(PRV))       #-- Number of Provinces
```

There are `r nrow(address_clean %>% distinct(PRV))`: too many to plot.

Let's explore the region for each client (`REGION`):

```{r Distribution REGION Address}
df3_dist_region<- address_clean                   %>%
  group_by(REGION)                                %>%  #-- Grouped By REGION
  summarize(TOT_ADDRESS = n_distinct(ID_ADDRESS)) %>%  #-- Number of Total Client
  mutate(PERCENT = TOT_ADDRESS/sum(TOT_ADDRESS))  %>%  #-- Percent
  arrange(desc(PERCENT))                               #-- Sorted by PERCENT 

knitr::kable(head(df3_dist_region))

ggplot(data = df3_dist_region, aes(x = REGION,
                                   y = TOT_ADDRESS)) + #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +                       #-- Bar Plot
  theme_minimal()                                      #-- ggplot Theme
```

## 4. Privacy

We check for duplicate rows:

```{r Duplicate Privacy}
knitr::kable(privacy_clean %>%
               summarize(TOT_ID_CLIs = n_distinct(ID_CLI), #-- Total Client
               TOT_ROWs = n()))                            #-- No Duplicated rows
```

Let's manipulate the privacy dataset:

```{r Manipulation Privacy}
privacy_clean <- privacy_clean                       %>%
  mutate(FLAG_PRIVACY_1 = as.factor(FLAG_PRIVACY_1)) %>% #-- Factorized
  mutate(FLAG_PRIVACY_2 = as.factor(FLAG_PRIVACY_2)) %>% #-- Factorized
  mutate(FLAG_DIRECT_MKT = as.factor(FLAG_DIRECT_MKT))   #-- Factorized
```

### Explanatory Analysis

Let's explore the first privacy method (FLAG_PRIVACY_1):

```{r Dsitribution FLAG_PRIVACY_1 Address}
df4_dist_privacy1<- privacy_clean                            %>%
                      group_by(FLAG_PRIVACY_1)               %>% #-- Grouped By FLAG_PRIVACY_1
                      summarize(TOT_ID = n_distinct(ID_CLI)) %>% #-- Number of Total Client
                      mutate(PERCENT = TOT_ID/sum(TOT_ID))   %>% #-- Percent
                      arrange(desc(PERCENT))                     #-- Sorted by PERCENT 

knitr::kable(df4_dist_privacy1)

ggplot(data = df4_dist_privacy1,
       aes(x = FLAG_PRIVACY_1,
           y = TOT_ID)) +        #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  theme_minimal()                #-- ggplot Theme
```

Let's explore the second privacy method (FLAG_PRIVACY_2):

```{r Dsitribution FLAG_PRIVACY_2 Address}
df4_dist_privacy2<- privacy_clean                            %>%
                      group_by(FLAG_PRIVACY_2)               %>% #-- Grouped By FLAG_PRIVACY_1
                      summarize(TOT_ID = n_distinct(ID_CLI)) %>% #-- Number of Total Client
                      mutate(PERCENT = TOT_ID/sum(TOT_ID))   %>% #-- Percent
                      arrange(desc(PERCENT))                     #-- Sorted by PERCENT 

knitr::kable(df4_dist_privacy2)

ggplot(data = df4_dist_privacy2,
       aes(x = FLAG_PRIVACY_2,
           y = TOT_ID)) +        #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  theme_minimal()                #-- ggplot Theme
```

Let's explore the privacy MTK method (FLAG_DIRECT_MKT):

```{r Dsitribution FLAG_DIRECT_MKT Address}
df4_dist_privacy_mkt<- privacy_clean                             %>%
                          group_by(FLAG_DIRECT_MKT)              %>% #-- Grouped By FLAG_PRIVACY_1
                          summarize(TOT_ID = n_distinct(ID_CLI)) %>% #-- Number of Total Client
                          mutate(PERCENT = TOT_ID/sum(TOT_ID))   %>% #-- Percent
                          arrange(desc(PERCENT))                     #-- Sorted by PERCENT 

knitr::kable(df4_dist_privacy_mkt)

ggplot(data = df4_dist_privacy_mkt,
       aes(x = FLAG_DIRECT_MKT,
           y = TOT_ID)) +        #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  theme_minimal()                #-- ggplot Theme
```

## 5. Campaign Category

The campaign channel (`CHANNEL_CAMP`) is not importants: let's remove it.

```{r Manipulation Campaign Category}
campaign_category_clean <- campaign_category_clean %>%
                             select(-CHANNEL_CAMP)
```

## 6. Campaign Event

Let's format dates and times:

```{r Manipulation Campaign Event}
campaign_event_clean <- campaign_event_clean                                                    %>%
                              mutate(EVENT_DATETIME = as.POSIXct(EVENT_DATE,
                                                                 format = "%Y-%m-%dT%H:%M:%S")) %>% #-- Formatting Dates
                              mutate(EVENT_HOUR = hour(EVENT_DATETIME))                         %>% #-- Hours
                              mutate(EVENT_DATE = as.Date(EVENT_DATETIME))                          #-- As Date
```

### Reshaping

We remap the event type values `E` (*ERROR*) and `B` [*BOUNCE*] into a level `F` [*FAILURE*]:

```{r Remap TYP_EVENT Campaign Event}
campaign_event_clean <- campaign_event_clean %>%
                          mutate(TYP_EVENT = as.factor(if_else(TYP_EVENT == "E" | TYP_EVENT == "B",
                                                               "F", as.character(TYP_EVENT)))) %>%  #-- Remapping Values
                          left_join(campaign_category_clean,
                                    by = "ID_CAMP")                                                 #-- Left Join

```

Also, is necessary to organize the data adding to each sending event the corresponding opens/clicks/fails:

```{r Sends Campaign Event}
df_sends <- campaign_event_clean             %>%
              filter(TYP_EVENT == "S")       %>% #-- Filter Sends
              select(-TYP_EVENT)             %>%
              select(ID_EVENT_S = ID_EVENT,
                     ID_CLI,
                     ID_CAMP,
                     TYP_CAMP,
                     ID_DELIVERY,
                     SEND_DATE = EVENT_DATE) %>% #-- Variable Selection
              as.data.frame()
```

There could be multiple opens of the same communication:

1. Count the open events;
2. Consider explicitely only the first open.

```{r Open Campaign Event}
df_opens_prep <- campaign_event_clean              %>%
                    filter(TYP_EVENT == "V")       %>% #-- Filter Sends
                    select(-TYP_EVENT)             %>%
                    select(ID_EVENT_O = ID_EVENT,
                           ID_CLI,
                           ID_CAMP,
                           TYP_CAMP,
                           ID_DELIVERY,
                           OPEN_DATETIME = EVENT_DATETIME,
                           OPEN_DATE = EVENT_DATE) %>% #-- Variable Selection
                    as.data.frame()

total_opens <- df_opens_prep                                  %>%
                  group_by(ID_CLI,
                           ID_CAMP,
                           ID_DELIVERY)                       %>% #-- Grouped By Client, Campaign & Delivery
                  summarize(NUM_OPENs = n_distinct(ID_EVENT_O))   #-- Number of Total Mail Opened
                
                df_opens <- df_opens_prep                     %>%
                  left_join(total_opens,
                            by = c("ID_CLI",
                                   "ID_CAMP",
                                   "ID_DELIVERY"))            %>% #-- Left join with total_opens
                  group_by(ID_CLI,
                           ID_CAMP,
                           ID_DELIVERY)                       %>% #-- Grouped By Client, Campaign & Delivery
                  filter(OPEN_DATETIME == min(OPEN_DATETIME)) %>% #-- Filter By Min Mail Opened
                  filter(row_number() == 1)                   %>%
                  ungroup()                                   %>%
                  as.data.frame()
```

There could be multiple clicks of the same communication:

1. Count the click events;
2. Consider explicitely only the first click.

```{r Clicks Campaign Event}
df_clicks_prep <- campaign_event_clean                              %>%
                      filter(TYP_EVENT == "C")                      %>% #-- Filter By Clicks
                      select(-TYP_EVENT)                            %>%
                      select(ID_EVENT_C = ID_EVENT,
                             ID_CLI,
                             ID_CAMP,
                             TYP_CAMP,
                             ID_DELIVERY,
                             CLICK_DATETIME = EVENT_DATETIME,
                             CLICK_DATE = EVENT_DATE)                   #-- Variable Selection
                    
total_clicks <- df_clicks_prep                                    %>%
                    group_by(ID_CLI,
                             ID_CAMP,
                             ID_DELIVERY)                         %>% #-- Grouped By Client, Campaign & Delivery
                    summarize(NUM_CLICKs = n_distinct(ID_EVENT_C))    #-- Number of Total Clicks

df_clicks <- df_clicks_prep                                   %>%
                left_join(total_clicks,
                          by = c("ID_CLI",
                                 "ID_CAMP",
                                 "ID_DELIVERY"))              %>% #-- Left Join with total_clicks
                group_by(ID_CLI,
                         ID_CAMP,
                         ID_DELIVERY)                         %>% #-- Grouped By Client, Campaign & Delivery
                filter(CLICK_DATETIME == min(CLICK_DATETIME)) %>% #-- Filter By Min Mail Clicked
                filter(row_number() == 1) %>%
                ungroup()                                     %>%
                as.data.frame()
```

```{r Fails Campaign Event}
df_fails <- campaign_event_clean              %>%
              filter(TYP_EVENT == "F")                    %>% #-- Filter By Fail
              select(-TYP_EVENT)                          %>%
              select(ID_EVENT_F = ID_EVENT,
                     ID_CLI,
                     ID_CAMP,
                     TYP_CAMP,
                     ID_DELIVERY,
                     FAIL_DATETIME = EVENT_DATETIME,
                     FAIL_DATE = EVENT_DATE)              %>% #-- Variable Selection
              group_by(ID_CLI,
                       ID_CAMP,
                       ID_DELIVERY)                       %>% #-- Grouped By Client, Campaign & Delivery
              filter(FAIL_DATETIME == min(FAIL_DATETIME)) %>% #-- Filter By Min Mail Failed
              filter(row_number() == 1)                   %>%
              ungroup()                                   %>%
              as.data.frame()
```

We combine them in one variable:

```{r Combine Campaign Event}
campaign_event_clean_final <- df_sends                     %>%
                                left_join(df_opens,
                                          by = c("ID_CLI",
                                                 "ID_CAMP",
                                                 "ID_DELIVERY",
                                                 "TYP_CAMP"))                            %>% #-- Left Join with df_opens
                                filter(is.na(OPEN_DATE) | SEND_DATE <= OPEN_DATE)        %>%
                                left_join(df_clicks,
                                          by = c("ID_CLI",
                                                 "ID_CAMP",
                                                 "ID_DELIVERY",
                                                 "TYP_CAMP"))                            %>% #-- Left Join with df_clicks
                                filter(is.na(CLICK_DATE) | OPEN_DATE <= CLICK_DATE)      %>%
                                left_join(df_fails,
                                          by = c("ID_CLI",
                                                 "ID_CAMP",
                                                 "ID_DELIVERY",
                                                 "TYP_CAMP"))                            %>% #-- Left Join with df_fails
                                filter(is.na(FAIL_DATE) | SEND_DATE <= FAIL_DATE)        %>% 
                                mutate(OPENED = !is.na(ID_EVENT_O))                      %>% #-- Opened
                                mutate(CLICKED = !is.na(ID_EVENT_C))                     %>% #-- Clicked
                                mutate(FAILED = !is.na(ID_EVENT_F))                      %>% #-- Failed
                                mutate(DAYS_TO_OPEN = as.integer(OPEN_DATE - SEND_DATE)) %>%
                                select(ID_EVENT_S,
                                       ID_CLI,
                                       ID_CAMP,
                                       TYP_CAMP,
                                       ID_DELIVERY,
                                       SEND_DATE,
                                       OPENED,
                                       OPEN_DATE,
                                       DAYS_TO_OPEN,
                                       NUM_OPENs,
                                       CLICKED,
                                       CLICK_DATE,
                                       NUM_CLICKs,
                                       FAILED)                                                #-- Variable Selection
```

### Explanatory Analysis

```{r Summaries Campaign Event}
df6_overview <- campaign_event_clean_final %>% 
                  summarize(MIN_DATE = min(SEND_DATE),           #-- Min Date
                            MAX_DATE = max(SEND_DATE),           #-- Max Date
                            TOT_EVENTs = n_distinct(ID_EVENT_S), #-- Number of Total Events
                            TOT_CLIs = n_distinct(ID_CLI))       #-- Number of Total Clicks

knitr::kable(df6_overview)
```

Below there's a general overview about the type of campaign variable (`TYP_CAMP`):

```{r Distribution TYP_CAMP Campaign Event}
df6_overviewbytyp <- campaign_event_clean_final %>%
                        group_by(TYP_CAMP)      %>%                    #-- Grouped By TYP_CAMP
                        summarize(MIN_DATE = min(SEND_DATE),           #-- Min Date
                                  MAX_DATE = max(SEND_DATE),           #-- Max Date
                                  TOT_EVENTs = n_distinct(ID_EVENT_S), #-- Number of Total Events
                                  TOT_CLIs = n_distinct(ID_CLI))       #-- Number of Total Clicks

knitr::kable(df6_overviewbytyp)

ggplot(data = df6_overviewbytyp,
       aes(x = TYP_CAMP,
           y = TOT_EVENTs)) +    #-- Dataset to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  theme_minimal()                #-- ggplot Theme
```

Below there's a general overview about the variable `OPENED`:

```{r Distribution OPENED Campaign Event}
df6_dist_opened <- campaign_event_clean_final                        %>%
                      group_by(OPENED)                               %>%
                      summarize(TOT_EVENTs = n_distinct(ID_EVENT_S),              #-- Number of Total Events
                                TOT_CLIs = n_distinct(ID_CLI))       %>%          #-- Number of Total Clicks                  
                      mutate(TYP_CAMP = 'ALL')                       %>%
                      mutate(PERCENT_EVENTs = TOT_EVENTs/df6_overview$TOT_EVENTs, #-- Percent
                             PERCENT_CLIs = TOT_CLIs/df6_overview$TOT_CLIs)       #-- Percent

knitr::kable(df6_dist_opened)

ggplot(data = df6_dist_opened,
         aes(fill = OPENED,
             x = TYP_CAMP,
             y = TOT_EVENTs)) +   #-- Data to Plot
    geom_bar(stat = "identity",
             position = "fill") + #-- Bar Plot
    theme_minimal()               #-- ggplot Theme
```

Below there's a general overview about the type of campaign variable opened (`TYP_CAMP`):

```{r Distribution OPENED by TYP_CAMP Campaign Event}
df6_dist_openedbytyp <- campaign_event_clean_final                         %>%
                          group_by(TYP_CAMP, OPENED)                       %>% #-- Grouped By TYP_CAMP & OPENED
                          summarize(TOT_EVENTs = n_distinct(ID_EVENT_S),
                                    TOT_CLIs = n_distinct(ID_CLI))         %>% #-- Number of Total Events & Client
                          left_join(df6_overviewbytyp %>%
                                      select(TYP_CAMP,
                                             ALL_TOT_EVENTs = TOT_EVENTs,
                                             ALL_TOT_CLIs = TOT_CLIs),
                                    by = 'TYP_CAMP')                       %>% #-- Left Join with df6_overviewbytyp
                          mutate(PERCENT_EVENTs = TOT_EVENTs/ALL_TOT_EVENTs,
                                 PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs)     %>% #-- Percent
                          select(TYP_CAMP,
                                 OPENED,
                                 TOT_EVENTs,
                                 TOT_CLIs,
                                 PERCENT_EVENTs,
                                 PERCENT_CLIs)                                 #-- Variable Selection

knitr::kable(df6_dist_openedbytyp)

ggplot(data = df6_dist_openedbytyp,
       aes(fill = OPENED,
           x = TYP_CAMP,
           y = TOT_EVENTs)) +       #-- Data to Plot
  geom_bar(stat = "identity") +     #-- Bar Plot
  theme_minimal()                   #-- ggplot Theme

ggplot(data = df6_dist_openedbytyp,
       aes(fill = OPENED,
           x = TYP_CAMP,
           y = TOT_EVENTs)) +       #-- Data to Plot
  geom_bar(position = "fill",
           stat = "identity") +     #-- Bar Plot
  theme_minimal()                   #-- ggplot Theme
```

Below there's a general overview about the days after an email is opened (`DAYS_TO_OPEN`):

```{r Distribution DAYS_TO_OPEN Campaign Event}
df6_dist_daystoopen <- campaign_event_clean_final                                 %>%
                          filter(OPENED)                                          %>%
                          group_by(ID_CLI)                                        %>% #-- Grouped By ID_CLI
                          summarize(AVG_DAYS_TO_OPEN = floor(mean(DAYS_TO_OPEN))) %>% #-- Mean Day to Open
                          ungroup()                                               %>%
                          group_by(AVG_DAYS_TO_OPEN)                              %>% #-- Grouped By AVG_DAYS_TO_OPEN
                          summarize(TOT_CLIs = n_distinct(ID_CLI))

knitr::kable(df6_dist_daystoopen)

ggplot(data = df6_dist_daystoopen %>%
         filter(AVG_DAYS_TO_OPEN < 14),
       aes(x = AVG_DAYS_TO_OPEN,
           y = TOT_CLIs)) +             #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") +        #-- Bar Plot
  theme_minimal()                       #-- ggplot Theme
```

Below there's a general overview about the days after an email is opened (`DAYS_TO_OPEN`):

```{r Distribution AVG_DAYS_TO_OPEN Campaign Event}
df6_dist_daystoopen_vs_cumulate <- df6_dist_daystoopen          %>%
                                      arrange(AVG_DAYS_TO_OPEN) %>%
                                      mutate(PERCENT_COVERED = cumsum(TOT_CLIs)/sum(TOT_CLIs))

ggplot(data = df6_dist_daystoopen_vs_cumulate %>%
         filter(AVG_DAYS_TO_OPEN < 14),
       aes(x = AVG_DAYS_TO_OPEN,
           y = PERCENT_COVERED)) +           #-- Data to Plot
  geom_line() +
  geom_point() +                             #-- Bar Plot
  scale_x_continuous(breaks = seq(0, 14, 2),
                     minor_breaks = 0:14) +  #-- Scale X
  theme_minimal()                            #-- ggplot Theme
```

Below there's a general overview about the days after an email is clicked:

```{r Distribution CLICKED by TYP_CAMP Campaign Event}
df6_dist_clickedbytyp <- campaign_event_clean_final                      %>%
                            group_by(TYP_CAMP,
                                     CLICKED)                            %>% #-- Grouped By TYP_CAMP & CLICKED
                            summarize(TOT_EVENTs = n_distinct(ID_EVENT_S),
                                      TOT_CLIs = n_distinct(ID_CLI))     %>% #-- Number of Total Event & Client
                            left_join(df6_overviewbytyp                  %>%
                                        select(TYP_CAMP,
                                               ALL_TOT_EVENTs = TOT_EVENTs,
                                               ALL_TOT_CLIs = TOT_CLIs),
                                      by='TYP_CAMP')                     %>% #-- Left Join with df6_overviewbytyp
                            mutate(PERCENT_EVENTs = TOT_EVENTs/ALL_TOT_EVENTs,
                                   PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs) %>% #-- Percent
                            select(TYP_CAMP,
                                   CLICKED,
                                   TOT_EVENTs,
                                   TOT_CLIs,
                                   PERCENT_EVENTs,
                                   PERCENT_CLIs)                             #-- Variable Selection

knitr::kable(df6_dist_clickedbytyp)

ggplot(data = df6_dist_clickedbytyp,
       aes(fill = CLICKED,
           x = TYP_CAMP,
           y = TOT_EVENTs)) +   #-- Data to Plot
  geom_bar(stat = "identity") + #-- Bar Plot
  theme_minimal()               #-- ggplot Theme

ggplot(data = df6_dist_clickedbytyp,
       aes(fill = CLICKED,
             x = TYP_CAMP,
             y = TOT_EVENTs)) + #-- Data to Plot
  geom_bar(position = "fill",
           stat = "identity") + #-- Bar Plot
  theme_minimal()               #-- ggplot Theme

```

Below there's a general overview about the days after an email is failed:

```{r Distribution Campaign Event}
df6_dist_failedbytyp <- campaign_event_clean_final                     %>%
                          group_by(TYP_CAMP,
                                   FAILED)                             %>% #-- Grouped By TYP_CAMP & FAILED
                          summarize(TOT_EVENTs = n_distinct(ID_EVENT_S),
                                    TOT_CLIs = n_distinct(ID_CLI))     %>% #-- Number of Total Client & Events
                          left_join(df6_overviewbytyp                  %>%
                                      select(TYP_CAMP,
                                             ALL_TOT_EVENTs = TOT_EVENTs,
                                             ALL_TOT_CLIs = TOT_CLIs),
                                    by = 'TYP_CAMP')                   %>% #-- Left Join with df6_overviewbytyp
                          mutate(PERCENT_EVENTs = TOT_EVENTs/ALL_TOT_EVENTs,
                                 PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs) %>% #-- Percent
                          select(TYP_CAMP,
                                 FAILED,
                                 TOT_EVENTs,
                                 TOT_CLIs,
                                 PERCENT_EVENTs,
                                 PERCENT_CLIs)                             #-- Variable Selection

knitr::kable(df6_dist_failedbytyp)

ggplot(data = df6_dist_failedbytyp,
       aes(fill = FAILED,
           x = TYP_CAMP,
           y = TOT_EVENTs)) +   #-- Data to Plot
  geom_bar(stat = "identity") + #-- Bar Plot
  theme_minimal()               #-- ggplot Theme

ggplot(data = df6_dist_failedbytyp,
       aes(fill = FAILED,
           x = TYP_CAMP,
           y = TOT_EVENTs)) +   #-- Data to Plot
  geom_bar(position = "fill",   
           stat = "identity") + #-- Bar Plot
  theme_minimal()               #-- ggplot Theme
```

Below there's a general overview about the number of clicks opening an email:

```{r Distribution NUM_OPENs Campaign Event}
df6_dist_numopens <- campaign_event_clean_final                  %>%
                      group_by(NUM_OPENs)                        %>% #-- Grouped By NUM_OPENs
                      summarize(TOT_ID = n_distinct(ID_EVENT_S)) %>% #-- Number of Total Event
                      mutate(PERCENT = TOT_ID/sum(TOT_ID))       %>% #-- Percent
                      arrange(desc(PERCENT))                         #-- Sort By Percent

df6_dist_numopens

ggplot(data = df6_dist_numopens,
       aes(x = NUM_OPENs,
           y = TOT_ID)) +        #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  xlim(0, 15)+                   #-- X Limit
  theme_minimal()                #-- ggplot Theme
```

Below there's a general overview about the number of clicks in the email:

```{r Distribution NUM_CLICKs Campaign Event}
df6_dist_numclicks <- campaign_event_clean_final                   %>%
                        group_by(NUM_CLICKs)                       %>% #-- Grouped By NUM_CLICKs
                        summarize(TOT_ID = n_distinct(ID_EVENT_S)) %>% #-- Number of Total Event
                        mutate(PERCENT = TOT_ID/sum(TOT_ID))       %>% #-- Percent
                        arrange(desc(PERCENT))                         #-- Sort By Percent


knitr::kable(df6_dist_numclicks)

ggplot(data = df6_dist_numclicks,
       aes(x = NUM_CLICKs,
           y = TOT_ID)) +        #-- Data to Plot
  geom_bar(stat = "identity",
           fill = "steelblue") + #-- Bar Plot
  xlim(0, 15) +                  #-- X Limit
  theme_minimal()                #-- ggplot Theme
```

## 7. Purchase and Refund

Let's format dates and times and factorize `DIREZIONE`:

```{r Manipulation Tickets}
tickets_clean <- tickets_clean %>%
                    mutate(TIC_DATETIME = as.POSIXct(DATETIME,
                                                     format = "%Y-%m-%dT%H%M%S")) %>%
                    mutate(TIC_HOUR = hour(TIC_DATETIME)) %>%
                    mutate(TIC_DATE = as.Date(TIC_DATETIME)) %>%
                    select(-DATETIME) %>%
                    mutate(DIREZIONE = as.factor(DIREZIONE)) %>% #-- Factorization
                    mutate(COD_REPARTO = as.factor(COD_REPARTO))
```

### Reshaping

```{r Reshaping Tickets}
tickets_clean <- tickets_clean %>%
                    #-- adding day characterization
                    mutate(TIC_DATE_WEEKDAY = wday(TIC_DATE)) %>%
                    mutate(TIC_DATE_HOLIDAY = isHoliday("Italy", TIC_DATE)) %>%
                    mutate(TIC_DATE_TYP = case_when(
                      (TIC_DATE_WEEKDAY %in% c(6,7)) ~ "weekend",
                      (TIC_DATE_HOLIDAY == TRUE) ~ "holiday",
                      (TIC_DATE_WEEKDAY < 7) ~ "weekday",
                      TRUE ~ "other"))
```

### Explanatory Analysis

Let's take a general overiew about the `TIC_DATE` variable:

```{r Distribution TIC_DATE Tickets}
df7_overview <- tickets_clean %>% 
                  summarize(MIN_DATE = min(TIC_DATE),
                            MAX_DATE = max(TIC_DATE),
                            TOT_TICs = n_distinct(ID_SCONTRINO),
                            TOT_CLIs = n_distinct(ID_CLI))

knitr::kable(df7_overview)
```

Let's take a general overiew about the `DIREZIONE` variable:

```{r Distribution DIREZIONE Tickets}
df7_dist_direction <- tickets_clean %>%
  group_by(DIREZIONE) %>%
  summarize(TOT_TICs = n_distinct(ID_SCONTRINO),
            TOT_CLIs = n_distinct(ID_CLI)) %>%
  mutate(PERCENT_TICs = TOT_TICs/df7_overview$TOT_TICs,
         PERCENT_CLIs = TOT_CLIs/df7_overview$TOT_CLIs)

knitr::kable(df7_dist_direction)
```

Let's take a general overiew about the `TIC_HOURS` variable:

```{r Distribution TIC_HOURS Tickets}
df7_dist_hour <- tickets_clean %>%
  group_by(TIC_HOUR, DIREZIONE) %>%
  summarize(TOT_TICs = n_distinct(ID_SCONTRINO),
            TOT_CLIs = n_distinct(ID_CLI)) %>%
  left_join(df7_dist_direction %>%
              select(DIREZIONE,
                     ALL_TOT_TICs = TOT_TICs,
                     ALL_TOT_CLIs = TOT_CLIs),
            by = 'DIREZIONE') %>%
  mutate(PERCENT_TICs = TOT_TICs/ALL_TOT_TICs,
         PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs) %>%
  select(-ALL_TOT_TICs,
         -ALL_TOT_CLIs)

knitr::kable(df7_dist_hour)

#-- TIC_HOURS
ggplot(data = df7_dist_hour,
       aes(fill = DIREZIONE,
           x = TIC_HOUR,
           y = TOT_TICs)) +
  geom_bar(stat = "identity") +
  theme_minimal()

#-- TIC_HOURS Percent
ggplot(data = df7_dist_hour,
       aes(fill = DIREZIONE,
           x = TIC_HOUR,
           y = TOT_TICs)) +
  geom_bar(stat = "identity",
           position = "fill" ) +
  theme_minimal()
```

Let's take a general overiew about the `COD_REPARTO` variable:

```{r Distribution COD_REPARTO Tickets}
df7_dist_dep <- tickets_clean %>%
  group_by(COD_REPARTO, DIREZIONE) %>%
  summarize(TOT_TICs = n_distinct(ID_SCONTRINO),
            TOT_CLIs = n_distinct(ID_CLI)) %>%
  left_join(df7_dist_direction %>%
              select(DIREZIONE,
                     ALL_TOT_TICs = TOT_TICs,
                     ALL_TOT_CLIs = TOT_CLIs),
            by = 'DIREZIONE') %>%
  mutate(PERCENT_TICs = TOT_TICs/ALL_TOT_TICs,
         PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs) %>%
    select(-ALL_TOT_TICs, -ALL_TOT_CLIs)
    
knitr::kable(df7_dist_dep)

#-- COD_REPARTO
ggplot(data = df7_dist_dep,
       aes(fill = DIREZIONE,
           x = COD_REPARTO,
           y = TOT_TICs)) +
  geom_bar(stat = "identity") +
  theme_minimal()

#-- COD_REPARTO Percent
ggplot(data = df7_dist_dep,
       aes(fill = DIREZIONE,
           x = COD_REPARTO,
           y = TOT_TICs)) +
  geom_bar(stat = "identity",
           position = "fill" ) +
  theme_minimal()
```

Let's take a general overiew about the `TIC_DATE_TYP` variable:

```{r Distribution TIC_DATE_TYP Tickets}
df7_dist_datetyp <- tickets_clean %>%
  group_by(TIC_DATE_TYP, DIREZIONE) %>%
  summarize(TOT_TICs = n_distinct(ID_SCONTRINO),
            TOT_CLIs = n_distinct(ID_CLI)) %>%
  left_join(df7_dist_direction %>%
              select(DIREZIONE,
                     ALL_TOT_TICs = TOT_TICs,
                     ALL_TOT_CLIs = TOT_CLIs),
            by = 'DIREZIONE') %>%
  mutate(PERCENT_TICs = TOT_TICs/ALL_TOT_TICs,
         PERCENT_CLIs = TOT_CLIs/ALL_TOT_CLIs) %>%
  select(-ALL_TOT_TICs, -ALL_TOT_CLIs)

knitr::kable(df7_dist_datetyp)

#-- TIC_DATE_TYP
ggplot(data = df7_dist_datetyp,
       aes(fill = DIREZIONE,
           x = TIC_DATE_TYP,
           y = TOT_TICs)) +
 geom_bar(stat = "identity") +
 theme_minimal()

#-- TIC_DATE_TYP Percent
ggplot(data = df7_dist_datetyp,
       aes(fill = DIREZIONE,
           x = TIC_DATE_TYP,
           y = TOT_TICs)) +
  geom_bar(stat = "identity",
           position = "fill" ) +
  theme_minimal()
```

Let's take a general overiew about the `IMPORTO_LORDO` variable:

```{r Distribution IMPORTO_LORDO Tickets}
df7_dist_importosconto <- tickets_clean %>%
  group_by(ID_SCONTRINO, DIREZIONE) %>%
  summarize(IMPORTO_LORDO = sum(IMPORTO_LORDO),
            SCONTO = sum(SCONTO)) %>%
  ungroup() %>%
  as.data.frame()

df7_dist_avgimportosconto <- df7_dist_importosconto %>%
  group_by(DIREZIONE) %>%
  summarize(AVG_IMPORTO_LORDO = mean(IMPORTO_LORDO),
            AVG_SCONTO = mean(SCONTO))

knitr::kable(df7_dist_avgimportosconto)

#-- IMPORTO_LORDO
ggplot(data = df7_dist_importosconto %>%
         filter((IMPORTO_LORDO > -1000) & (IMPORTO_LORDO < 1000)),
       aes(color = DIREZIONE,
           x = IMPORTO_LORDO)) +
  geom_histogram(binwidth = 10,
                 fill = "white",
                 alpha = 0.5) +
  theme_minimal()

#-- IMPORTO_LORDO Percent
ggplot(data = df7_dist_importosconto %>%
         filter((SCONTO > -250) & (IMPORTO_LORDO < 250)),
       aes(color = DIREZIONE, x = SCONTO)) +
  geom_histogram(binwidth = 10,
                 fill = "white",
                 alpha = 0.5) +
  theme_minimal()
```

```{r Distribution Average IMPORTO_LORDO Tickets}
df7_dist_importosconto_cod_rep <- tickets_clean %>%
  group_by(COD_REPARTO, DIREZIONE) %>%
  summarize(IMPORTO_LORDO = sum(IMPORTO_LORDO),
            SCONTO = sum(SCONTO)) %>%
  ungroup() %>%
  as.data.frame()

knitr::kable(df7_dist_importosconto_cod_rep)

#-- IMPORTO_LORDO
ggplot(data = df7_dist_importosconto_cod_rep,
       aes(fill = DIREZIONE,
           x = COD_REPARTO,
           y = IMPORTO_LORDO)) +
 geom_bar(stat = "identity") +
 theme_minimal()

#-- SCONTO
ggplot(data = df7_dist_importosconto_cod_rep,
       aes(fill = DIREZIONE,
           x = COD_REPARTO,
           y = SCONTO)) +
 geom_bar(stat = "identity") +
 theme_minimal()
```

```{r Distribution ID_ARTICOLO}
tickets_clean$ID_ARTICOLO <- as.factor(tickets_clean$ID_ARTICOLO)

tickets_clean

df7_dist_id_articolo <- tickets_clean %>%
  filter(DIREZIONE == 1) %>%
  group_by(ID_ARTICOLO) %>%
  summarize(NUM_VENDITE = n_distinct(ID_SCONTRINO)) %>%
  ungroup() %>%
  as.data.frame() %>%
  arrange(desc(NUM_VENDITE))

knitr::kable(head(df7_dist_id_articolo))
```

```{r Distribution Average ID_CLIENTE Tickets}
df7_dist_importosconto_id_cli <- tickets_clean %>%
  filter(DIREZIONE == 1) %>%
  group_by(ID_CLI) %>%
  summarize(IMPORTO_LORDO = sum(IMPORTO_LORDO),
            SCONTO = sum(SCONTO)) %>%
  ungroup() %>%
  as.data.frame() %>%
  arrange(desc(IMPORTO_LORDO))

knitr::kable(head(df7_dist_importosconto_id_cli))
```

```{r}
df7_dist_tot_purch <- tickets_clean %>%
                        filter(DIREZIONE == 1) %>%
                        group_by(ID_CLI) %>%
                        summarise(TOT_PURCHASE = n_distinct(ID_SCONTRINO)) %>%
                        arrange(desc(TOT_PURCHASE))

knitr::kable(head(df7_dist_tot_purch))
```

```{r}
df7_dist_recency <- tickets_clean %>%
  filter(DIREZIONE == 1) %>%
  group_by(ID_CLI) %>%
  summarise(LAST_PURCHASE_DATE = max(TIC_DATE))

df7_dist_recency$RECENCY <- difftime(as.Date("30/04/2019", format = "%d/%m/%Y"),
                                     df7_dist_recency$LAST_PURCHASE_DATE, units = "days")

knitr::kable(head(df7_dist_recency))
```

# RFM Model

```{r}
rfm_study_period <- tickets_clean %>%
  filter(TIC_DATE > as.Date("01/01/2019", format = "%d/%m/%Y")) #-- Active 

knitr::kable(head(rfm_study_period))
```

```{r}
rfm_recency <- rfm_study_period %>%
  filter(DIREZIONE == 1) %>%
  group_by(ID_CLI) %>%
  summarise(LAST_PURCHASE_DATE = max(TIC_DATE))

rfm_recency$RECENCY <- difftime(as.Date("30/04/2019", format = "%d/%m/%Y"),          #-- Recency
                                     rfm_recency$LAST_PURCHASE_DATE, units = "days")

knitr::kable(head(rfm_recency))
```

```{r}
rfm_frequency <- rfm_study_period %>%
                    filter(DIREZIONE == 1) %>%
                    group_by(ID_CLI) %>%
                    summarise(TOT_PURCHASE = n_distinct(ID_SCONTRINO)) %>%
                    arrange(desc(TOT_PURCHASE))

knitr::kable(head(rfm_frequency))
```

```{r}
rfm_monetary <- rfm_study_period %>%
  filter(DIREZIONE == 1) %>%
  group_by(ID_CLI) %>%
  summarize(IMPORTO_LORDO = sum(IMPORTO_LORDO),
            SCONTO = sum(SCONTO),
            SPESA = IMPORTO_LORDO - SCONTO) %>%
  ungroup() %>%
  as.data.frame() %>%
  arrange(desc(IMPORTO_LORDO))

knitr::kable(head(rfm_monetary))
```

```{r}
# 0-2 -> "low"
# 3-4 -> "medium"
# 5+ -> "high"

rfm_frequency <- within(rfm_frequency,
                   FREQ_CLASS <- cut(rfm_frequency$TOT_PURCHASE,
                                              breaks = c(0, 2, 5, 101),
                                              include.lowest=T,
                                     right = F,
                                     labels = c("low", "medium", "high")))
table(rfm_frequency$FREQ_CLASS)
```

```{r}
rfm_recency <- within(rfm_recency,
                   REC_CLASS <- cut(as.numeric(rfm_recency$RECENCY),
                                              breaks = quantile(rfm_recency$RECENCY, probs=c(0, .25, .75, 1)),
                                              include.lowest=T,
                                     labels = c("low", "medium", "high")),
                   )

table(rfm_recency$REC_CLASS)
```

```{r}
rfm_monetary <- within(rfm_monetary,
                   MON_CLASS <- cut(rfm_monetary$SPESA,
                                              breaks = quantile(rfm_monetary$SPESA, probs=c(0, .25, .75, 1)),
                                              include.lowest=T,
                                     labels = c("low", "medium", "high")),
                   )

table(rfm_monetary$MON_CLASS)
```


```{r}
rfm <- merge(rfm_frequency, rfm_monetary, by = "ID_CLI")
rfm <- merge(rfm, rfm_recency, by = "ID_CLI")

rfm[,c("ID_CLI","REC_CLASS","FREQ_CLASS","MON_CLASS")]

save(rfm,rfm_monetary,rfm_recency,rfm_frequency, file = "rfm.RData")
```

```{r}
rfm$RF <- NA

for(i in c(1:nrow(rfm))){
  if(rfm$REC_CLASS[i] == "low" && rfm$FREQ_CLASS[i] == "low") rfm$RF[i] <- "One-Timer"
  if(rfm$REC_CLASS[i] == "medium" && rfm$FREQ_CLASS[i] == "low") rfm$RF[i] <- "One-Timer"
  if(rfm$REC_CLASS[i] == "high" && rfm$FREQ_CLASS[i] == "low") rfm$RF[i] <- "Leaving"
  if(rfm$REC_CLASS[i] == "low" && rfm$FREQ_CLASS[i] == "medium") rfm$RF[i] <- "Engaged"
  if(rfm$REC_CLASS[i] == "medium" && rfm$FREQ_CLASS[i] == "medium") rfm$RF[i] <- "Engaged"
  if(rfm$REC_CLASS[i] == "high" && rfm$FREQ_CLASS[i] == "medium") rfm$RF[i] <- "Leaving"
  if(rfm$REC_CLASS[i] == "low" && rfm$FREQ_CLASS[i] == "high") rfm$RF[i] <- "Top"
  if(rfm$REC_CLASS[i] == "medium" && rfm$FREQ_CLASS[i] == "high") rfm$RF[i] <- "Top"
  if(rfm$REC_CLASS[i] == "high" && rfm$FREQ_CLASS[i] == "high") rfm$RF[i] <- "Leaving Top"
}

table(rfm$RF)
```